Eagle: Making multiple-locus association mapping on a genome-wide scale routine

### 2018-08-01 paper.tex
1August refining v2

The application of regularisation methods to association mapping is appealing because the strength of association between all the SNPs and 
trait can be measured simultaneously. 
 The downside is that it is challenging to calculate  
the statistical significance of a SNP effect, analytically. We instead calculate significance, empirically, via stability selection (see below). 
%% Permutation has been suggest as an empirical solution to this 
%% limitation \cite{shen2013novel} 
%% but we have instead opted for stability selection (see below).

MLMM is close in philosophy to Eagle. It too is based on  building the 
best  linear mixed model via forward selection using the extBIC criterion. 
However, there are several differences between the two methods.
 MLMM does not make use of dimension reduction, which gives it a larger  computational footprint than Eagle. 
  It also builds its model differently to Eagle. Eagle uses a score statistic in which to identify SNPs to enter the model. 
  MLMM uses the statistical significance of a SNP when it is treated as a fixed effect in the model. 
 This necessitates fitting a separate linear mixed model for each unidentified SNP. Both are R packages but much of Eagle is written in 
 C++ code. There are significant differences in run time and memory usage (see Results).

R2VIM is different to the other four methods in that it is a non-parametric model-free approach. It implements  
random forests and a new way of measuring the importance of a SNP.  
In random forests, the worth of a predictor is measured, empirically, by calculating its importance score. 
It is from these importance scores that SNPs can be ordered in terms of their strength of association with a trait. 
The difficulty is in knowing what proportion of the 
highest ordered SNPs are in true association. R2VIM addresses this by performing multiple parallel runs. Each run uses a different 
random seed, giving slightly different results for the random forest analyses. Within a run, a relative importance score ($f$) is calculated for 
each SNP. The greater $f$ is than one, the more evidence there is that the SNP is in association with the trait.  A SNP-trait association is 
declared if $f$ for the SNP is above some threshold value across all parallel runs. The difficulty is that the relationship between 
this threshold value and the chance of finding false positives is unknown.  The threshold value could be found via permutation but 
this would be computationally challenging for all but the smallest of association studies.

>>>>

MLMM is closest in philosophy to Eagle. It too is based on building the best model via forward selection, 
within  a linear mixed model framework, using the 
extBIC. However, there are differences between the two methods.  MLMM does not make use of dimension reduction. Also, it builds its "best" 
model differently to Eagle.  Eagle uses a score statistic in which to identify SNPs to enter the model.
MLMM uses the statistical significance of the SNPs. This involves fitting a separate linear mixed model for each unidentified SNP. 
Both are R packages but there is a significant difference in computational performance (see Results).

R2VIM is different to the other four methods in that it is a non-parametric model-free approach. It implements  
random forests and a new way of measuring the importance of a SNP.  
Here, multiple parallel runs are performed. Within a run, a relative importance score is calculated for each SNP and 
calculated by dividing the importance score for a SNP by a run's minimum importance score. 
Only those SNPs with relative importance scores above a certain threshold across all the runs are deemed to be significant. 
The relationship between threshold value and false positive rate is unknown. The threshold could be found empirically via permutation 
but the computation cost is high.

 The greater the score is than one, 
the stronger the evidence is that the SNP is in association with the trait.  Only those SNPs with scores above a certain threshold are 
deemed as being significant. The difficulty is that the relationship between this threshold and false positive rate is unknown. The threshold could be 
found empirically, via permutation but this is computationally demanding.

In random forests, the worth of a predictor is measured, empirically, by calculating its importance score. 
It is from these importance scores that SNPs can be ordered in terms of their strength of association with a trait. 
The difficulty is in knowing what proportion of the 
highest ordered SNPs are in true association. R2VIM addresses this by performing multiple parallel runs. Each run uses a different 
random seed, giving slightly different results for the random forest analyses. Within a run, a relative importance score ($f$) is calculated for 
each SNP. The greater $f$ is than one, the more evidence there is that the SNP is in association with the trait.  A SNP-trait association is 
declared if $f$ for the SNP is above some threshold value across all parallel runs. The difficulty is that the relationship between 
this threshold value and the chance of finding false positives is unknown.  The threshold value could be found via permutation but 
this would be computationally challenging for all but the smallest of association studies.
